hbm_paper_results

### 2024-02-21 paper_lbcc_results.Rmd
parse results, plots

Combat-GAM Centiles

```{r}
gam_centile_df <- fread(file=paste0(der_path, "/cf.gam-batch.study_predictions.csv"))
dim(gam_centile_df)
```

Combat-LS Centiles
```{r}
gamlss_centile_df <- fread(file=paste0(der_path, "/cf.gamlss-batch.study_predictions.csv"))
dim(gamlss_centile_df)
```

@@

```{r}
deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
# skimr::skim(deltas_df)
cent_deltas_df <- deltas_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
```

@@

Sample
```{r}
sample_boxplot <- ggplot(cent_deltas_df, aes(y=reorder(study, -age_years), x=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank()) +
  labs(color="Sex", y = "Study")

#overall sample
full_boxplot <- 
  cent_deltas_df %>%
  mutate(ylab = "   Full Sample") %>%
  ggplot(aes(y=ylab, x=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  # theme(plot.margin = margin(0,0,0,0, 'cm')) +
  # scale_y_discrete(labels=c("1" = "Full Sample")) +
  labs(color="Sex", x = "Age (Years)", y="")

sample_barplot <- ggplot(cent_deltas_df) +
  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=reorder(study, -age_years), fill=sex)) + 
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  labs(color="Sex")

@@

```{r}
delta_summary.df <- cent_deltas_df_long %>%
  group_by(sex, study) %>%
  summarise(mean_delta = mean(delta),
            mean_abs.delta = mean(abs(delta)),
            max_abs.delta=max(abs(delta)),
            sex_n = n())

abs_delta_summary.df <- cent_deltas_df %>%
  pivot_longer(cols=starts_with("delta."), names_to="pheno", names_prefix = "delta.", values_to = "delta") %>%
  mutate(abs.delta = abs(delta)) %>%
  group_by(study) %>%
  summarise(study_mean_abs.delta=mean(abs.delta),
            study_mean_delta=mean(delta),
            mean_age=mean(age_days),
            mean_age_yrs=mean(age_days)/365.25,
            study_max_abs.delta=max(abs.delta),
            study_n=n())
```

```{r}
delta_sum.df <- full_join(delta_summary.df, abs_delta_summary.df) %>%
  mutate(sex_prop=sex_n/study_n,
         majority = ifelse(sex_prop > .5, 1, 0))

delta_sum.df_piv <- delta_sum.df %>%
  pivot_wider(id_cols=c(study, study_n, mean_age, mean_age_yrs, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta, max_abs.delta))

#Merge back female prop
to_merge <- delta_sum.df_piv %>%
  dplyr::select(study, sex_prop_Female)

delta_sum.df2 <- full_join(delta_sum.df, to_merge)
```

@@

```{r}
delta_sum.df2 %>%
  ggplot(aes(x=sex_prop_Female, y=mean_delta, color=sex)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")

delta_sum.df2 %>%
  ggplot(aes(x=sex_prop_Female, y=mean_delta, color=study, shape=sex)) +
  geom_point() +
  geom_smooth(aes(color=sex, linetype=sex), method="lm")

delta_sum.df2 %>%
  ggplot(aes(x=sex_prop_Female, y=mean_delta, color=sex)) +
  geom_point() +
  geom_smooth(aes(color=sex, linetype=sex), method="lm")

delta_sum.df %>%
  ggplot(aes(x=sex, y=mean_delta, color=study, fill=study)) +
  geom_count(aes(size=sex_prop), alpha=0.5) +
  geom_line(aes(group=study))

delta_sum.df %>%
  ggplot(aes(x=sex, y=mean_delta, color=study, fill=study)) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=study))

delta_sum.df %>%
  ggplot(aes(x=sex_prop, y=mean_abs.delta, color=mean_age_yrs)) +
  geom_count(aes(size=sex_n)) +
  geom_smooth(method="lm") +
  facet_wrap(~sex)

delta_sum.df_piv %>%
  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta, color=mean_age_yrs)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")
```

#Across-Feature Mean Results

```{r}
# age
summary(lm(study_mean_abs.delta ~ mean_age_yrs, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ mean_age_yrs, data=delta_sum.df_piv))
```

@@

```{r}
delta_sum.df %>%
  ggplot(aes(x=mean_age_yrs, y=mean_abs.delta, color=sex)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")

age_effects.plt <- delta_sum.df_piv %>%
  ggplot(aes(x=mean_age_yrs, y=(study_mean_abs.delta*100))) +
  geom_count(aes(size=study_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  labs(size="Study N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores")
```

are studies w younger subjects changing more because there are fewer of them?
check for association between study size and age
```{r}
summary(lm(mean_age_yrs ~ study_n, data=delta_sum.df_piv))
fit <- lm((study_mean_abs.delta*100) ~ mean_age_yrs + study_n, data=delta_sum.df_piv)
summary(fit)
sjPlot::plot_model(fit, type = "pred", terms = "mean_age_yrs")

summary(lm((study_mean_abs.delta*100) ~ mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_piv))
```

@@

```{r}
# %female
summary(lm(study_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ sex_prop_Female, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ sex_prop_Female + study_n, data=delta_sum.df_piv)) #add study size covariate
summary(lm(study_mean_delta ~ sex_prop_Female + study_n + mean_age, data=delta_sum.df_piv)) #add study size & age covariates
```

```{r}
delta_sum.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100), color=study, fill=study)) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=study))

sex_effects.plt <- delta_sum.df_piv %>%
  ggplot(aes(x=sex_prop_Female, y=(study_mean_delta*100))) +
  geom_count(aes(size=study_n, color=mean_age_yrs)) +
  geom_smooth(method="lm", color="#24693DFF") +
  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="Study N", x = "Proportion Female", y = "Mean Difference in Centile Scores")

delta_sum.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100))) +
  geom_boxplot()
```

```{r}
# age*%female
summary(lm(study_mean_abs.delta ~ mean_age_yrs*sex_prop_Female, data=delta_sum.df_piv))
summary(lm(study_mean_delta ~ mean_age_yrs*sex_prop_Female, data=delta_sum.df_piv))
```

#Featurewise Results

Linear models
```{r}
cent_delta_lms.df <- fread(file=paste0(der_path, "/batch.study_pred_cent_delta_lms.csv"))
n_unique(cent_delta_lms.df$pheno)
```

summarise findings

```{r}
cent_delta_lms.df %>%
  dplyr::filter(term=="sexMale") %>%
  dplyr::summarize(n_sig_regions = sum(sig_fdr),
                   n_regions=nrow(.),
              mean_est = mean(estimate)) %>%
    ungroup() %>%
    as.data.frame()
```

```{r}
cent_delta_lms.df %>%
  dplyr::filter(term=="age_yrs") %>%
  dplyr::summarize(n_sig_regions = sum(sig_fdr),
                   n_regions=nrow(.),
              mean_est = mean(estimate)) %>%
    ungroup() %>%
    as.data.frame()
```

```{r}
cent_delta_lms.df %>%
  dplyr::filter(term=="sexMale:age_yrs") %>%
  dplyr::summarize(n_sig_regions = sum(sig_fdr),
                   n_regions=nrow(.),
              mean_est = mean(estimate)) %>%
    ungroup() %>%
    as.data.frame()
```

```{r}
sig.cent_delta_lms.df <- cent_delta_lms.df %>%
  dplyr::filter(sig_fdr==TRUE)
n_unique(sig.cent_delta_lms.df$pheno)

table(sig.cent_delta_lms.df$term, sig.cent_delta_lms.df$pheno)

max(sig.cent_delta_lms.df$estimate)*100
min(sig.cent_delta_lms.df$estimate)*100
```

@@

Plot betas (*100 for centile space) on brain!

```{r}
cent_delta_lms.df_toplt <- cent_delta_lms.df %>%
   mutate(pheno_cat = as.factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    label = sub("_[^_]*_", "_", pheno)) #for plotting
```

```{r}
term_names <- c(sexMale = "Sex", age_yrs = "Age", 'sexMale:age_yrs' = "Sex*Age")

#Regional
regional.delta_plt <- cent_delta_lms.df_toplt %>%
  dplyr::filter(pheno_cat != "Global Volume") %>%
  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
                                     TRUE ~ estimate*100)) %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = estimate)) +
  facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.11, 0.11), breaks=c(-0.1, 0, 0.1))

#Global
global.delta_plt <- cent_delta_lms.df_toplt %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
                                     TRUE ~ estimate*100),
         tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA))) %>%
  dplyr::select(!any_of(c("label", "region", "hemi"))) %>% #drop cols corresponding to dk regions
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             side = c("coronal"),
             aes(fill = estimate)) +
  facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names)), switch = 'y') + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.11, 0.11), breaks=c(-0.1, 0, 0.1))
```

@@

df_gam_feat_aov <- data.frame()
for(pheno in gam.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gam_feat_aov <- rbind(df_gam_feat_aov, df)
}

@@

```{r}
gam_centile_df.clean %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=WMV)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

summary(lm(WMV ~ study + age_days + sex, data = gam_centile_df.clean))
```

@@

```{r}
gam_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

@@

```{r}
summary(lm(mean_cent ~ study, data=gam_centile_df.mean))

summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
pairwise.t.test(gam_centile_df.mean$mean_cent, gam_centile_df.mean$study, p.adjust.method = "fdr")

mean_cent_gam.lm <- lm(mean_cent ~ study + age_days + sex, data=gam_centile_df.mean)

gtsummary::tbl_regression(mean_cent_gam.lm) %>%
  gtsummary::add_global_p()
```

@@

```{r}
gam.feat_to_test.z <- paste0(gam.feat_to_test, ".z")
df_gam_feat_aov.z <- data.frame()
for(pheno in gam.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gam_feat_aov.z <- rbind(df_gam_feat_aov.z, df)
}

@@

```{r}
gam_z_df.mean <- gam_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(gam_z_df.mean) == n_unique(gam_z_df.clean$participant)
```

@@

```{r}
summary(lm(mean_z ~ study, data=gam_z_df.mean))

summary(aov(mean_z ~ study, data=gam_z_df.mean))
pairwise.t.test(gam_z_df.mean$mean_z, gam_z_df.mean$study, p.adjust.method = "fdr")

mean_z_gam.lm <- lm(mean_z ~ study + age_days + sex, data=gam_z_df.mean)
summary(mean_z_gam.lm)
```

@@

df_gamlss_feat_aov <- data.frame()
for(pheno in feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
}

@@

```{r}
gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  theme(axis.text.x = element_text(angle=90))

gamlss_centile_df.mean %>%
  ggplot() +
  geom_violin(aes(x=reorder(study, age_days), y=mean_cent), scale="count") + 
  theme(axis.text.x = element_text(angle=90))

@@

```{r}
summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))

summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
pairwise.t.test(gamlss_centile_df.mean$mean_cent, gamlss_centile_df.mean$study, p.adjust.method = "fdr")

mean_cent_gamlss.lm <- lm(mean_cent ~ study + age_days + sex + fs_version, data=gamlss_centile_df.mean)
summary(mean_cent_gamlss.lm)

mean_cent_gamlss.gam <- gam(mean_cent ~ study + s(log_age) + sex, data=gamlss_centile_df.mean)
summary(mean_cent_gamlss.gam)

mean_cent_gam.gam <- gam(mean_cent ~ study + s(log_age) + sex, data=gam_centile_df.mean)
summary(mean_cent_gam.gam)

gtsummary::tbl_regression(mean_cent_gamlss.lm) %>%
  gtsummary::add_global_p()
```

@@

```{r}
gamlss.feat_to_test.z <- paste0(feat_to_test, ".z")
df_gamlss_feat_aov.z <- data.frame()
for(pheno in gamlss.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
}

@@

```{r}
gamlss_z_df.mean <- gamlss_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(gamlss_z_df.mean) == n_unique(gamlss_z_df.clean$participant)
```

@@

```{r}
summary(lm(mean_z ~ study, data=gamlss_z_df.mean))

summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
pairwise.t.test(gamlss_z_df.mean$mean_z, gamlss_z_df.mean$study, p.adjust.method = "fdr")

mean_cent_gamlss.lm <- lm(mean_z ~ study + age_days + sex, data=gamlss_z_df.mean)
summary(mean_cent_gamlss.lm)
```

@@

```{r}
# study demographics
fig5 <- ggarrange(
  ggarrange(
ggarrange(sample_boxplot, sample_barplot,
          full_boxplot, full_barplot,
          nrow=2, ncol=2, 
          # align = "hv", 
          common.legend = TRUE,
          legend = "bottom",
          widths=c(2, 1),
          heights=c(6, 1)),

>>>>

Combat-GAM Centiles - site-level batch correction

```{r}
gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.site_predictions.csv"))
dim(gam_centile_df)
```

Combat-LS Centiles
```{r}
gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.site_predictions.csv"))
dim(gamlss_centile_df)
```

@@

```{r}
deltas_df <- fread(file=paste0(der_path, "/batch.site_pred_deltas.csv"))
# skimr::skim(deltas_df)
cent_deltas_df <- deltas_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

tbl.df <- as.data.frame(table(cent_deltas_df$site, cent_deltas_df$study))
```

@@

Sample plot
```{r}
#reorder studies by age
age.df <- cent_deltas_df %>%
  group_by(study) %>%
  summarize(mean_age = mean(age_years)) %>%
  arrange(mean_age)

sample_boxplot <- cent_deltas_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
ggplot(aes(y=interaction(site,study, sep = "&"), x=age_years)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text=element_blank(), ggh4x.axis.nesttext.y = element_text(size=6)) +
  labs(color="Sex", y = "Site")

#overall sample
full_boxplot <- 
  cent_deltas_df %>%
  mutate(ylab = "  Full \n  Sample") %>%
  ggplot(aes(y=ylab, x=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  # theme(plot.margin = margin(0,0,0,0, 'cm')) +
  # scale_y_discrete(labels=c("1" = "Full Sample")) +
  labs(color="Sex", x = "Age (Years)", y="")

sample_barplot <- cent_deltas_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
  ggplot() +
  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=interaction(site,study, sep = "&"), fill=sex)) +
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  labs(color="Sex")

@@

```{r}
delta_n.sex <- cent_deltas_df_long %>%
  mutate(abs.delta = abs(delta)) %>%
  group_by(sex, study, site) %>%
  summarise(sex_n = n(),
            mean_abs.delta=mean(abs.delta),
            mean_delta=mean(delta),
            mean_age_yrs=mean(age_days)/365.25) #get number males and females in each site

#get site-level stats
delta_summary.df <- cent_deltas_df_long %>%
  mutate(abs.delta = abs(delta)) %>%
  group_by(study, site) %>%
  summarise(site_mean_abs.delta=mean(abs.delta),
            site_mean_delta=mean(delta),
            site_mean_age=mean(age_days),
            site_mean_age_yrs=mean(age_days)/365.25,
            site_max_abs.delta=max(abs.delta),
            site_n=n())

delta_n.sex %>%
  dplyr::filter(site=="abide2_BNI_1")
```

```{r}
delta_sum.by.sex.df <- full_join(delta_summary.df, delta_n.sex) %>%
  mutate(sex_prop=sex_n/site_n,
         majority = ifelse(sex_prop > .5, 1, 0))

delta_sum.by.sex.df_piv <- delta_sum.by.sex.df %>%
  pivot_wider(id_cols=c(study, site, site_n, site_mean_age, site_mean_age_yrs, site_mean_delta, site_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta))

#Merge back female prop back into overall site-level stats
to_merge <- delta_sum.by.sex.df_piv %>%
  dplyr::select(study, site, sex_prop_Female) %>%
  mutate_at(vars(sex_prop_Female), ~replace_na(., 0)) #set prop Female to 0 if no females in study

delta_sum.df_merged <- full_join(delta_summary.df, to_merge)
```

@@

```{r}
delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")

delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=mean_delta, color=site, fill=site)) +
  geom_count(aes(size=sex_prop), alpha=0.5) +
  geom_line(aes(group=site)) +
  theme(legend.position="none")

delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=mean_delta, color=as.factor(majority), fill=as.factor(majority))) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=site), color="darkgray")

delta_sum.by.sex.df %>%
  ggplot(aes(x=sex_prop, y=mean_abs.delta, color=mean_age_yrs)) +
  geom_count(aes(size=sex_n)) +
  geom_smooth(method="lm") +
  facet_wrap(~sex)

delta_sum.by.sex.df_piv %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta, color=study)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")
```

look at max abs. delta across sites
```{r}
delta_sum.df_merged %>%
  ggplot() +
  geom_histogram(aes(x=site_max_abs.delta*100))

delta_sum.df_merged %>%
  ggplot(aes(y=site_max_abs.delta*100, x=site_n)) +
  geom_point() +
  geom_smooth(method="lm")

delta_sum.df_merged %>%
  ggplot(aes(y=site_max_abs.delta*100, x=site_mean_age_yrs)) +
  geom_point() +
  geom_smooth(method="lm")

delta_sum.df_merged %>%
  ggplot(aes(y=site_max_abs.delta*100, x=sex_prop_Female)) +
  geom_point() +
  geom_smooth(method="lm")
```

##Feature-wise
do deltas differ by feature?
```{r}
cent_deltas_df_long2 <- cent_deltas_df_long %>%
   mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

cent_deltas_df_long2 %>%
  ggplot() +
  geom_histogram(aes(x=delta, fill=pheno), position="identity", alpha=0.3) +
  theme(legend.position="none")

cent_deltas_df_long2 %>%
  ggplot() +
  geom_histogram(aes(x=delta*100, fill=pheno_cat, color=pheno_cat), position="identity", alpha=0.7, bins=60)

cent_deltas_df_long2 %>%
  ggplot() +
  geom_histogram(aes(x=delta*100, fill=study, color=study), position="identity", alpha=0.3, bins=60) +
  facet_wrap(~pheno_cat, scales="free") +
  theme(legend.position="none")
```

#Site-Level Mean Results

@@

```{r}
# age
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ site_mean_age_yrs, data=delta_sum.df_merged))

#forward selection for site_mean_abs.delta
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
full.delta.age.fit <- lm((site_mean_abs.delta*100) ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
sjPlot::plot_model(full.delta.age.fit, type = "pred", terms = "site_mean_age_yrs")
```

```{r}
delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=site_mean_abs.delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")

age_effects.plt <- delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
  geom_count(aes(size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores")

# skim(delta_sum.df_piv) #some missing data! check on this!
```

are studies w older subjects changing more because there are fewer of them?
check for association between study size and age
```{r}
summary(lm(site_mean_age_yrs ~ site_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
```

@@

```{r}
# %female
summary(lm(site_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ sex_prop_Female + site_n, data=delta_sum.df_merged)) #add site size covariate
summary(lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged)) #add site size & age covariates
full.delta.sex.fit <- lm((100*site_mean_delta) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged) #scaling for plot

sjPlot::plot_model(full.delta.sex.fit, type = "pred", terms = "sex_prop_Female") 
```

```{r}
delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=site))

sex_effects.plt <- delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=(site_mean_delta*100))) +
  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
  geom_smooth(method="lm", color="#24693DFF") +
  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores")

delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100))) +
  geom_boxplot()
```

```{r}
# age*%female
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
```

@@

@@

df_gam_feat_aov <- data.frame()
for(pheno in gam.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["site"]]))
  df$pheno <- pheno
  df_gam_feat_aov <- rbind(df_gam_feat_aov, df)
}

@@

Test single pheno

```{r}
gam_centile_df.clean %>%
  ggplot() +
  geom_boxplot(aes(x=site, y=WMV)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

summary(lm(WMV ~ site + age_days + sex, data = gam_centile_df.clean))
```

@@

```{r}
gam_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=site, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

gam_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

@@

```{r}
summary(lm(mean_cent ~ site, data=gam_centile_df.mean))
summary(lm(mean_cent ~ study, data=gam_centile_df.mean))

summary(aov(mean_cent ~ site, data=gam_centile_df.mean))
```

success!

@@

```{r}
gam.feat_to_test.z <- paste0(gam.feat_to_test, ".z")
df_gam_feat_aov.z <- data.frame()
for(pheno in gam.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["site"]]))
  df$pheno <- pheno
  df_gam_feat_aov.z <- rbind(df_gam_feat_aov.z, df)
}

@@

```{r}
gam_z_df.mean <- gam_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(gam_z_df.mean) == n_unique(gam_z_df.clean$participant)
```

@@

```{r}
summary(lm(mean_z ~ site, data=gam_z_df.mean))
summary(lm(mean_z ~ study, data=gam_z_df.mean))

summary(aov(mean_z ~ study, data=gam_z_df.mean))
```

@@

df_gamlss_feat_aov <- data.frame()
for(pheno in feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["site"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
}

@@

```{r}
gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(site, age_days), y=mean_cent)) + 
  theme(axis.text.x = element_text(angle=90))

gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  theme(axis.text.x = element_text(angle=90))

@@

```{r}
summary(lm(mean_cent ~ site, data=gamlss_centile_df.mean))
summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))

summary(aov(mean_cent ~ site, data=gamlss_centile_df.mean))
```

success!

@@

```{r}
gamlss.feat_to_test.z <- paste0(feat_to_test, ".z")
df_gamlss_feat_aov.z <- data.frame()
for(pheno in gamlss.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["site"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
}

@@

```{r}
gamlss_z_df.mean <- gamlss_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site=unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(gamlss_z_df.mean) == n_unique(gamlss_z_df.clean$participant)
```

@@

```{r}
summary(lm(mean_z ~ site, data=gamlss_z_df.mean))
summary(lm(mean_z ~ study, data=gamlss_z_df.mean))

summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
```

@@

```{r}
# study demographics
fig5 <- ggarrange(
ggarrange(sample_boxplot, sample_barplot,
          full_boxplot, full_barplot,
          nrow=2, ncol=2, 
          # align = "hv", 
          common.legend = TRUE,
          legend = "bottom",
          widths=c(2, 1),
          heights=c(14, 1)),
# mean effects
ggarrange(age_effects.plt,
sex_effects.plt,
nrow=2, ncol=1,
align="hv",
labels=c("B", "C")),
common.legend=FALSE,
nrow=1, ncol=2,
labels=c("A", NA),
widths=c(1, 1.3)
)
